|

### 2021-05-19 pre-analysis-plan/freire-skarbek-lynching.tex
add pdfs and translations

---
title:  |
        | Pre-Analysis Plan:
        | Casting the First Stone: Understanding Attitudes towards Lynching in Brazil[^thanks]
author:
  - Danilo Freire[^freire]
  - David Skarbek[^skarbek]
date: \today
abstract: "Why do citizens support extrajudicial violence? While endorsement for vigilantism has significantly decreased in advanced democracies, lynchings remain tolerated throughout the developing world. In this pre-analysis plan, we propose three survey experiments to investigate (i) which criminal profiles citizens prefer for extralegal punishment, (ii) how individuals justify mob violence, and (iii) whether information provision reduces support for vigilantism. First, we run a conjoint experiment to identify criminal characteristics that are associated with higher likelihood of lynching. In our second experiment, we show respondents a short article about a real lynching episode and ask them whether they see the mob's motivations as legitimate. Lastly, we test whether providing information about the legal consequences of lynching, human rights guarantees, and the risk of retribution makes respondents less likely to endorse extralegal violence. We will run the experiments in Brazil, a country which has seen a sharp rise in vigilante attacks and currently experiences one lynching attempt per day. Our survey will be conducted online and include 2,000 participants."
abstractspacing: double
keywords: Brazil, crime, extralegal violence, lynching, vigilantism
fontsize: 12pt
margin: 2cm
urlcolor: darkblue
linkcolor: Mahogany
citecolor: Mahogany
spacing: double
papersize: a4paper
bibliography: references.bib
biblio-style: apalike
output:
  pdf_document:
    citation_package: natbib
    fig_caption: yes
    number_sections: yes
    keep_tex: yes
    toc: no
    toc_depth: 3
    template: template.latex
---

```{r, message=FALSE, warning=FALSE, echo=FALSE}
# If you need to install any package while knitting the document
r <- getOption("repos")
r["CRAN"] <- "https://cran.rstudio.com/"
options(repos = r)
if (!require("kableExtra")) {
    install.packages("kableExtra")
}
```

[^thanks]: We thank Rosario Aguilar, Louise Araújo, Miriam Golden, Umberto
Mignozzetti, Irfan Nooruddin, Catarina Roman, and Alex Scacco for their
valuable comments. This pre-analysis plan received IRB approval from Brown
University in October 2020 (Protocol 2009002803).

[^freire]: Postdoctoral Research Associate, The Political Theory Project, Brown
University, 8 Fones Alley, Providence, RI 02912,
[`danilofreire@gmail.com`](mailto:danilofreire@gmail.com),
<https://danilofreire.github.io>.

[^skarbek]: The Department of Political Science and the Political Theory
Project, Brown University, 8 Fones Alley, Providence, RI 02912, USA,
[`david_skarbek@brown.edu`](mailto:david_skarbek@brown.edu),
<http://davidskarbek.com>.

# Motivation
\label{sec:motivation}

@@

Why do citizens endorse extralegal punishment?[^definitions] Although mob
violence has decreased in advanced democracies, vigilantism remains widespread
in the developing world. In particular, Latin America has been severely
affected by a recent increase in lynchings. Motivated by growing numbers of
drug-related crime and by police ineffectiveness, citizen violence has surged
in the region [@mallen2014vigilantes]. Brazil provides a telling example. From
2011 to 2015, the country registered about 2,500 lynching episodes, and in 2015
alone 173 people were killed by angry mobs, an average of one execution every
two days [@barbara2015vigilantes; @oliveira2016mob]. According to José Martins,
who has studied lynchings in Brazil for more than 30 years, those figures are
not only the highest in the country's history, but also the highest in the
world [@pearson2018latam]. Victims are generally accused of petty theft, and
many are beaten to death for stealing mobile phones, cooking utensils, or pairs
of sandals [@barbara2015vigilantes]. Lynching perpetrators are often young men,
but they also include female teenagers, elderly women, and even members of
local police forces [@moura2017linchamentos; @oliveira2016mob].

[^definitions]: Here we consider extralegal violence, vigilantism, lynching,
and related terms as synonyms. We follow @moncada2017varieties [6] and define
vigilantism as "the collective use or threat of extralegal violence in response
to an alleged criminal act". While we focus here on vigilantism that directly
involves physical violence, we make no distinction between lethal and
non-lethal lynching episodes.

Recent research suggests that support for vigilantism in Latin America is
associated with exposure to violence [@garcia2019anger], untrustworthy law
enforcement [@zizumbo2017community], perceived insecurity [@ceobanu2011crime;
@godoy2004justice], and social inequality [@arias2010violent]. However,
approval for extrajudicial violence raises important questions about the
relationship between states and societies in the region
[@nivette2016institutional]. First, while vigilantism provides social order to
vulnerable populations, it also undermines the legitimacy of state institutions
[@schuberth2013challenging]. Lynching episodes reinforce the idea that
civilians should not wait for legal proceedings and that social control is
better exerted by the people themselves [@black1983crime]. Second, victim
characteristics greatly impact the likelihood and severity of lynchings. In
Brazil, a person is rarely lynched when they commit a crime against males, and
females that commit the same crimes as males also receive less
punishment.[^fapesp] Moreover, there is no significant social difference
between victims and perpetrators, nor are black Brazilians particularly
targeted by lynch mobs [@oliveira2016mob]. These dynamics pose a puzzle to
existing explanations for mob violence support and stand in sharp contrast with
the American experience [@smaangs2016doing; @wood2011lynching].

[^fapesp]: See <https://revistapesquisa.fapesp.br/en/mob-injustice>. Access:
June, 2020.

In this study, we endeavour to answer the following questions: 1) which victim
characteristics increase individual support for extralegal punishment?; 2) how
do individuals justify their approval for vigilantism?; and 3) can information
provision reduce support for mob violence? We design a set of three experiments
to tackle these issues. Our first experiment consists of a conjoint analysis in
which we show respondents different profiles of lynching victims and ask them
who they believe deserve punishment. In our second experiment, we test how
ineffective policing, slow criminal justice systems, and demand for hasher
legal punishment affect individual propensity to support mob violence. Finally,
we assess whether citizens become less supportive of vigilantism when they are
informed about the legal consequences of lynchings, constitutional human
rights, or the risk of _vendettas_ associated with mob violence. While in this
pre-analysis plan we present the three experiments as described above, we will
randomised their order in our survey to mitigate potential order bias.

# Experiment 01
\label{sub:exp01}

@@

* Lynchings are often used as social punishment in Brazil. Lynchings are cases
  in which three or more people physically attack or execute a suspected
  criminal in public. We are interested in knowing more about how Brazilians
  see these episodes. In the next five questions, please read the description
  of two possible lynching victims in Brazil and indicate in which case you
  believe the punishment is more justified. Even if you are not entirely sure,
  please select one of the cases.[^portuguese]

[^portuguese]: We will conduct the survey experiments in Portuguese. The
translation for the first experiment is as follows: Linchamentos são às vezes
usados como punição social no Brasil. Linchamentos são casos nos quais três ou
mais pessoas agridem fisicamente ou executam em público um suspeito de um
crime. Estamos interessados em saber mais sobre como os brasileiros vêem estes
episódios. Nas próximas cinco questões, por favor, leia a descrição de duas
possíveis vítimas de linchamento no Brasil e indique em quais delas você
acredita que a punição é mais justificada. Mesmo que você não tenha certeza,
por favor, escolha um dos casos.

## Sample
\label{sub:sample}

@@

## Estimation
\label{sub:estimation}

We will estimate our models with the `cregg` package [@leeper2018cregg] for the
R statistical language [@rstats2019]. We will report marginal means and average
marginal component effects (AMCEs) in our main analysis. However,
@leeper2018subgroup show that AMCEs can be misleading in subgroup comparisons
as model results are sensitive to the choice of reference categories in
interactions. In contrast, marginal means provide a clear description of
quantities of interest, in our case preferences towards lynching, while
allowing for easy comparisons between groups of respondents. So while we will
include both AMCEs and marginal means estimators in our article, we will
conduct our subgroup analysis using only the latter. We will cluster standard
errors by respondent.

## Hypotheses
\label{sec:hypotheses}

@@

1) Gender of crime perpetrator: male 
2) Age of crime perpetrator: teenager
3) Race of crime perpetrator: blacks
4) Residency of crime perpetrator: outsider
5) Offense: murder
6) Gender of victim: female 
7) Age of victim: child
8) Lynching perpetrators: the family of the victim

@@

## Heterogeneous Effects
\label{sub:het01}

@@

* Younger respondents believe that lynchings by the police are less justified.

* Female respondents believe that lynchings against females are less justified.

* Respondents of a particular race are less likely to choose individuals of
  the same race as lynching victims.

* More educated individuals show less support for lynchings against petty
  criminals.

* Political ideology has no effect on support for lynchings when the victim is
  female.

* Right-wingers show stronger support for lynchings against males than
  left-wingers.

* Respondents who were victimised in the last year are more likely to support
  lynchings by neighbours and the police.

# Experiment 02

@@

Research shows that police ineffectiveness frequently appears as a strong
predictor for vigilantism [@cruz2019determinants; @garcia2019anger]. The direct
result of the weakness of police institutions is that citizens decide to take
criminal matters "into their own hands", thus persecuting and punishing the
criminals by themselves. A recent statistic indicates that the police solve
only 10% of the homicides in Brazil, what lends support to the link between
weak law enforcement and lynchings [@pearson2018latam].

Another determinant of support for vigilante justice that is suggested in the
literature is the lack of trust in the justice system [@godoy2004justice;
@smith2019contradictions]. This is often due to long criminal proceedings,
which cause significant anxiety for the victims. In Brazil, the penal code
allows the accused to appeal each decision several times, so it can take
decades before a criminal case is closed [@sousa2005utilizaccao]. In this
respect, citizens do not believe that, even if the criminal is put to trial,
he/she will be punished in a timely matter. Note that although the police is
part of the criminal justice system, we separate the two institutions in our
experiment.

Lastly, we evaluate whether respondents think that the legal punishment
assigned to criminals is not proportional to the severity of their crimes. In
particular, we hope to gauge the demand for iron-fisted criminal justice in
Brazil. Although this treatment arm is related to the previous ones, it
addresses not the efficiency of the institutions, but their legitimacy
[@nivette2016institutional]. We believe that the demand for harsher laws is on
the rise in Brazil, and this may be a major reason why citizens justify their
support for mob violence. In fact, Brazilians are often vocal about their
preference for repressive legal punishment. In a recent article in _The Wall
Street Journal_, a bar owner justified the lynching of the local thug who
killed his son by saying that "_even if he had been put behind bars for 100
years it wouldn't have been enough to pay for all his crimes_"
[@pearson2018latam]. We hypothesise that many Brazilians also share this view.

The experiment consists of three treatment conditions and one control group.
Respondents will read an excerpt of a news article describing a real lynching
case. We have slightly edited the original text so that respondents have no
prior knowledge of the crime.[^jornal] The vignette for the control group
includes no information about the reasons behind the lynching. We will ask
respondents to show their level of support for mob violence using a 0-100
slider, where 0 means no support and 100 means full support. Respondents in
each of the three treatment arms will read the same piece, but with one
additional sentence explaining the motivations behind the lynching. The
vignettes are as follows:

[^jornal]: The original article is available at the following address:
<https://jr.jor.br/2020/05/01/homem-e-linchado-na-vila-progresso>.
Access: August 2020.

* *Control group*: A man was lynched last Friday in Jundiaí, São Paulo.
  According to the neighbours, he tried to break into a house but was
  immobilised and beaten by members of the community.[^control-exp02]

[^control-exp02]: In Portuguese: Um homem foi linchado na última sexta-feira em
Jundiaí, São Paulo. De acordo com vizinhos, ele tentou invadir uma residência
mas foi imobilizado e agredido por membros da comunidade.

* *Treatment 01 - Police ineffectiveness*: A man was lynched last Friday in
  Jundiaí, São Paulo. According to the neighbours, he tried to break into a
  house but was immobilised and beaten by members of the community. **One of
  the residents who took part in the lynching said they had beaten the suspect
  because "the police never patrols the area"**.[^treat01-exp02]

[^treat01-exp02]: In Portuguese: Um homem foi linchado na última sexta-feira em
Jundiaí, São Paulo. De acordo com vizinhos, ele tentou invadir uma residência
mas foi imobilizado e agredido por membros da comunidade. **Um dos moradores
envolvidos no linchamento disse que eles agrediram o suspeito porque "a polícia
nunca patrulha o local".**

* *Treatment 02 - Criminal justice ineffectiveness*: A man was lynched last
  Friday in Jundiaí, São Paulo. According to the neighbours, he tried to break
  into a house but was immobilised and beaten by members of the community.
  **One of the residents who took part in the lynching said they had beaten the
  suspect because "the judicial system is too slow and the perpetrator is on
  the street until the case is heard"**.[^treat02-exp02]

[^treat02-exp02]: In Portuguese: Um homem foi linchado na última sexta-feira em
Jundiaí, São Paulo. De acordo com vizinhos, ele tentou invadir uma residência
mas foi imobilizado e agredido por membros da comunidade. **Um dos moradores
envolvidos no linchamento disse que eles agrediram o suspeito porque "a justiça
é muito lenta e os criminosos ficam soltos até o julgamento"**.

* *Treatment 03 - Demand for harsher legal punishment*: A man was lynched last
  Friday in Jundiaí, São Paulo. According to the neighbours, he tried to break
  into a house but was immobilised and beaten by members of the community.
  **One of the residents who took part in the lynching said they had beaten the
  suspect because "the judicial punishment is not harsh
  enough"**.[^treat03-exp02]

[^treat03-exp02]: In Portuguese: Um homem foi linchado na última sexta-feira em
Jundiaí, São Paulo. De acordo com vizinhos, ele tentou invadir uma residência
mas foi imobilizado e agredido por membros da comunidade. **Um dos moradores
envolvidos no linchamento disse que eles agrediram o suspeito porque "a punição
da justiça não é dura o suficiente"**.

@@

* You will be shown a news article. Please read it carefully. After you read
  the article, we will ask you one question about it.[^vig01-exp02]

[^vig01-exp02]: In Portuguese: Uma notícia será apresentada para você. Por
favor, leia a notícia com atenção. Após você ler o artigo, faremos uma pergunta
sobre ele.

@@

* Would you think that the lynching was justified? Please use the slider below
  to indicate your opinion. For disagreement, use 0 to 49; for agreement,
  use 51 to 100. Please use 50 if you neither agree nor
  disagree.[^vig02-exp02]

[^vig02-exp02]: In Portuguese: Você acha que o linchamento foi correto? Por
favor, use a barra abaixo para indicar sua opinião. Para discordar, use de 0 a
49; para concordar, use de 51 a 100. Por favor, use 50 para não concordar nem
discordar.

## Sample

@@

## Estimation

@@

>>>>

% !TEX program = XeLaTeX
\documentclass[12pt,a4paper,]{article}

% Fonts
\usepackage{libertine}
% This is a nice mathfont, it fits well with Libertine
% Comment the line if you want to use a different one
\usepackage[libertine]{newtxmath}
% The default monofont 
% Again, comment the line below if you want to change it
\usepackage[scaled=.95]{inconsolata}

% Add the following packages to support kableExtra
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{etoolbox}
\usepackage{tocloft}

% Dots in toc
\renewcommand{\cftsecleader}{\cftdotfill{\cftdotsep}}

% Colours
\usepackage[usenames,dvipsnames]{xcolor}
\definecolor{darkblue}{rgb}{0.0,0.0,0.55}

% Spacing
\usepackage{setspace}

% Margin
\usepackage[margin=2cm]{geometry}

% Packages I've been using for different reasons...
\usepackage{hyperref}
\usepackage{dcolumn}
\usepackage{graphicx}
\usepackage{float}
\floatplacement{figure}{H}
\usepackage{pgf}
\usepackage{tikz}
\usetikzlibrary{arrows}
\usetikzlibrary{positioning}
\usepackage{mathtools}
\usepackage{caption}

% UK English
\usepackage[UKenglish]{babel}
\usepackage[UKenglish]{isodate}
\cleanlookdateon

% Penalties
\exhyphenpenalty=1000
\hyphenpenalty=1000
\widowpenalty=1000
\clubpenalty=1000

% Hypersetup
\hypersetup{
  linkcolor=Mahogany,
  citecolor=Mahogany,
  urlcolor=darkblue, 
  breaklinks=true, 
  colorlinks=true,
      pdfauthor={Danilo Freire; David Skarbek},
      pdfkeywords={Brazil, crime, extralegal violence, lynching, vigilantism},
  }

% If XeTex, LuaLaTeX, etc
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{amssymb,amsmath}
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}

% Language

% Bibliography
\usepackage{natbib}
\bibliographystyle{apalike}
\makeatletter
% Remove comma after author
\setcitestyle{aysep={}}
\patchcmd{\NAT@citex}
	  {\@citea\NAT@hyper@{%
		 \NAT@nmfmt{\NAT@nm}%
		 \hyper@natlinkbreak{\NAT@aysep\NAT@spacechar}{\@citeb\@extra@b@citeb}%
		 \NAT@date}}
	  {\@citea\NAT@nmfmt{\NAT@nm}%
	   \NAT@aysep\NAT@spacechar\NAT@hyper@{\NAT@date}}{}{}
	\patchcmd{\NAT@citex}
	  {\@citea\NAT@hyper@{%
		 \NAT@nmfmt{\NAT@nm}%
		 \hyper@natlinkbreak{\NAT@spacechar\NAT@@open\if*#1*\else#1\NAT@spacechar\fi}%
		   {\@citeb\@extra@b@citeb}%
		 \NAT@date}}
	  {\@citea\NAT@nmfmt{\NAT@nm}%
	   \NAT@spacechar\NAT@@open\if*#1*\else#1\NAT@spacechar\fi\NAT@hyper@{\NAT@date}}
	  {}{}
% Patch case where name and year are separated by aysep
\patchcmd{\NAT@citex}
  {\@citea\NAT@hyper@{%
     \NAT@nmfmt{\NAT@nm}%
     \hyper@natlinkbreak{\NAT@aysep\NAT@spacechar}{\@citeb\@extra@b@citeb}%
     \NAT@date}}
  {\@citea\NAT@nmfmt{\NAT@nm}%
   \NAT@aysep\NAT@spacechar\NAT@hyper@{\NAT@date}}{}{}
% Patch case where name and year are separated by opening bracket
\patchcmd{\NAT@citex}
  {\@citea\NAT@hyper@{%
     \NAT@nmfmt{\NAT@nm}%
     \hyper@natlinkbreak{\NAT@spacechar\NAT@@open\if*#1*\else#1\NAT@spacechar\fi}%
       {\@citeb\@extra@b@citeb}%
     \NAT@date}}
  {\@citea\NAT@nmfmt{\NAT@nm}%
   \NAT@spacechar\NAT@@open\if*#1*\else#1\NAT@spacechar\fi\NAT@hyper@{\NAT@date}}
  {}{}
\makeatother

% Listings

% Verbatim

% Tables

% Graphics

% Make links footnotes instead of hotlinks:
%  \setlength{\emergencystretch}{3em}  % prevent overfull lines
 \providecommand{\tightlist}{%
   \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

% Numbered sections
\setcounter{secnumdepth}{5}
% % % Redefines (sub)paragraphs to behave more like sections
% \ifx\paragraph\undefined\else
% \let\oldparagraph\paragraph
% \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
% \fi
% \ifx\subparagraph\undefined\else
% \let\oldsubparagraph\subparagraph
% \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
% \fi
% 
% Spacing
\doublespacing

% Title
\title{Pre-Analysis Plan:\\
Casting the First Stone: Understanding Attitudes towards Lynching in
Brazil\footnote{We thank Rosario Aguilar, Louise Araújo, Miriam Golden,
  Umberto Mignozzetti, Irfan Nooruddin, Catarina Roman, and Alex Scacco
  for their valuable comments. This pre-analysis plan received IRB
  approval from Brown University in October 2020 (Protocol 2009002803).}}

% Author
\author{Danilo Freire\footnote{Postdoctoral Research Associate, The Political
  Theory Project, Brown University, 8 Fones Alley, Providence, RI 02912,
  \href{mailto:danilofreire@gmail.com}{\texttt{danilofreire@gmail.com}},
  \url{https://danilofreire.github.io}.} \and David Skarbek\footnote{The Department of Political Science and the
  Political Theory Project, Brown University, 8 Fones Alley, Providence,
  RI 02912, USA,
  \href{mailto:david_skarbek@brown.edu}{\texttt{david\_skarbek@brown.edu}},
  \url{http://davidskarbek.com}.}}

% Date
\date{\today}

% Begin document
\begin{document}
\maketitle

% Abstract
\begin{abstract}
\doublespacing \noindent Why do citizens support extrajudicial violence? While endorsement for
vigilantism has significantly decreased in advanced democracies,
lynchings remain tolerated throughout the developing world. In this
pre-analysis plan, we propose three survey experiments to investigate
(i) which criminal profiles citizens prefer for extralegal punishment,
(ii) how individuals justify mob violence, and (iii) whether information
provision reduces support for vigilantism. First, we run a conjoint
experiment to identify criminal characteristics that are associated with
higher likelihood of lynching. In our second experiment, we show
respondents a short article about a real lynching episode and ask them
whether they see the mob's motivations as legitimate. Lastly, we test
whether providing information about the legal consequences of lynching,
human rights guarantees, and the risk of retribution makes respondents
less likely to endorse extralegal violence. We will run the experiments
in Brazil, a country which has seen a sharp rise in vigilante attacks
and currently experiences one lynching attempt per day. Our survey will
be conducted online and include 2,000 participants.
\vspace{.25cm}

\noindent \textbf{Keywords}: Brazil, crime, extralegal violence, lynching, vigilantism
\vspace{.25cm}

\end{abstract}

% Table of Contents
\newpage

\hypertarget{motivation}{%
\section{Motivation}\label{motivation}}

\label{sec:motivation}

@@

Why do citizens endorse extralegal punishment?\footnote{Here we consider
  extralegal violence, vigilantism, lynching, and related terms as
  synonyms. We follow \citet[6]{moncada2017varieties} and define
  vigilantism as ``the collective use or threat of extralegal violence
  in response to an alleged criminal act''. While we focus here on
  vigilantism that directly involves physical violence, we make no
  distinction between lethal and non-lethal lynching episodes.} Although
mob violence has decreased in advanced democracies, vigilantism remains
widespread in the developing world. In particular, Latin America has
been severely affected by a recent increase in lynchings. Motivated by
growing numbers of drug-related crime and by police ineffectiveness,
citizen violence has surged in the region \citep{mallen2014vigilantes}.
Brazil provides a telling example. From 2011 to 2015, the country
registered about 2,500 lynching episodes, and in 2015 alone 173 people
were killed by angry mobs, an average of one execution every two days
\citep{barbara2015vigilantes, oliveira2016mob}. According to José
Martins, who has studied lynchings in Brazil for more than 30 years,
those figures are not only the highest in the country's history, but
also the highest in the world \citep{pearson2018latam}. Victims are
generally accused of petty theft, and many are beaten to death for
stealing mobile phones, cooking utensils, or pairs of sandals
\citep{barbara2015vigilantes}. Lynching perpetrators are often young
men, but they also include female teenagers, elderly women, and even
members of local police forces
\citep{moura2017linchamentos, oliveira2016mob}.

Recent research suggests that support for vigilantism in Latin America
is associated with exposure to violence \citep{garcia2019anger},
untrustworthy law enforcement \citep{zizumbo2017community}, perceived
insecurity \citep{ceobanu2011crime, godoy2004justice}, and social
inequality \citep{arias2010violent}. However, approval for extrajudicial
violence raises important questions about the relationship between
states and societies in the region \citep{nivette2016institutional}.
First, while vigilantism provides social order to vulnerable
populations, it also undermines the legitimacy of state institutions
\citep{schuberth2013challenging}. Lynching episodes reinforce the idea
that civilians should not wait for legal proceedings and that social
control is better exerted by the people themselves
\citep{black1983crime}. Second, victim characteristics greatly impact
the likelihood and severity of lynchings. In Brazil, a person is rarely
lynched when they commit a crime against males, and females that commit
the same crimes as males also receive less punishment.\footnote{See
  \url{https://revistapesquisa.fapesp.br/en/mob-injustice}. Access:
  June, 2020.} Moreover, there is no significant social difference
between victims and perpetrators, nor are black Brazilians particularly
targeted by lynch mobs \citep{oliveira2016mob}. These dynamics pose a
puzzle to existing explanations for mob violence support and stand in
sharp contrast with the American experience
\citep{smaangs2016doing, wood2011lynching}.

In this study, we endeavour to answer the following questions: 1) which
victim characteristics increase individual support for extralegal
punishment?; 2) how do individuals justify their approval for
vigilantism?; and 3) can information provision reduce support for mob
violence? We design a set of three experiments to tackle these issues.
Our first experiment consists of a conjoint analysis in which we show
respondents different profiles of lynching victims and ask them who they
believe deserve punishment. In our second experiment, we test how
ineffective policing, slow criminal justice systems, and demand for
hasher legal punishment affect individual propensity to support mob
violence. Finally, we assess whether citizens become less supportive of
vigilantism when they are informed about the legal consequences of
lynchings, constitutional human rights, or the risk of \emph{vendettas}
associated with mob violence. While in this pre-analysis plan we present
the three experiments as described above, we will randomised their order
in our survey to mitigate potential order bias.

\hypertarget{experiment-01}{%
\section{Experiment 01}\label{experiment-01}}

\label{sub:exp01}

@@

\begin{itemize}
\tightlist
\item
  Lynchings are often used as social punishment in Brazil. Lynchings are
  cases in which three or more people physically attack or execute a
  suspected criminal in public. We are interested in knowing more about
  how Brazilians see these episodes. In the next five questions, please
  read the description of two possible lynching victims in Brazil and
  indicate in which case you believe the punishment is more justified.
  Even if you are not entirely sure, please select one of the
  cases.\footnote{We will conduct the survey experiments in Portuguese.
    The translation for the first experiment is as follows: Linchamentos
    são às vezes usados como punição social no Brasil. Linchamentos são
    casos nos quais três ou mais pessoas agridem fisicamente ou executam
    em público um suspeito de um crime. Estamos interessados em saber
    mais sobre como os brasileiros vêem estes episódios. Nas próximas
    cinco questões, por favor, leia a descrição de duas possíveis
    vítimas de linchamento no Brasil e indique em quais delas você
    acredita que a punição é mais justificada. Mesmo que você não tenha
    certeza, por favor, escolha um dos casos.}
\end{itemize}

\hypertarget{sample}{%
\subsection{Sample}\label{sample}}

\label{sub:sample}

@@

\hypertarget{estimation}{%
\subsection{Estimation}\label{estimation}}

\label{sub:estimation}

We will estimate our models with the \texttt{cregg} package
\citep{leeper2018cregg} for the R statistical language
\citep{rstats2019}. We will report marginal means and average marginal
component effects (AMCEs) in our main analysis. However,
\citet{leeper2018subgroup} show that AMCEs can be misleading in subgroup
comparisons as model results are sensitive to the choice of reference
categories in interactions. In contrast, marginal means provide a clear
description of quantities of interest, in our case preferences towards
lynching, while allowing for easy comparisons between groups of
respondents. So while we will include both AMCEs and marginal means
estimators in our article, we will conduct our subgroup analysis using
only the latter. We will cluster standard errors by respondent.

\hypertarget{hypotheses}{%
\subsection{Hypotheses}\label{hypotheses}}

\label{sec:hypotheses}

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi})}
\tightlist
\item
  Gender of crime perpetrator: male
\item
  Age of crime perpetrator: teenager
\item
  Race of crime perpetrator: blacks
\item
  Residency of crime perpetrator: outsider
\item
  Offense: murder
\item
  Gender of victim: female
\item
  Age of victim: child
\item
  Lynching perpetrators: the family of the victim
\end{enumerate}

@@

\hypertarget{heterogeneous-effects}{%
\subsection{Heterogeneous Effects}\label{heterogeneous-effects}}

\label{sub:het01}

@@

\begin{itemize}
\item
  Younger respondents believe that lynchings by the police are less
  justified.
\item
  Female respondents believe that lynchings against females are less
  justified.
\item
  Respondents of a particular race are less likely to choose individuals
  of the same race as lynching victims.
\item
  More educated individuals show less support for lynchings against
  petty criminals.
\item
  Political ideology has no effect on support for lynchings when the
  victim is female.
\item
  Right-wingers show stronger support for lynchings against males than
  left-wingers.
\item
  Respondents who were victimised in the last year are more likely to
  support lynchings by neighbours and the police.
\end{itemize}

\hypertarget{experiment-02}{%
\section{Experiment 02}\label{experiment-02}}

@@

Research shows that police ineffectiveness frequently appears as a
strong predictor for vigilantism
\citep{cruz2019determinants, garcia2019anger}. The direct result of the
weakness of police institutions is that citizens decide to take criminal
matters ``into their own hands'', thus persecuting and punishing the
criminals by themselves. A recent statistic indicates that the police
solve only 10\% of the homicides in Brazil, what lends support to the
link between weak law enforcement and lynchings
\citep{pearson2018latam}.

Another determinant of support for vigilante justice that is suggested
in the literature is the lack of trust in the justice system
\citep{godoy2004justice, smith2019contradictions}. This is often due to
long criminal proceedings, which cause significant anxiety for the
victims. In Brazil, the penal code allows the accused to appeal each
decision several times, so it can take decades before a criminal case is
closed \citep{sousa2005utilizaccao}. In this respect, citizens do not
believe that, even if the criminal is put to trial, he/she will be
punished in a timely matter. Note that although the police is part of
the criminal justice system, we separate the two institutions in our
experiment.

Lastly, we evaluate whether respondents think that the legal punishment
assigned to criminals is not proportional to the severity of their
crimes. In particular, we hope to gauge the demand for iron-fisted
criminal justice in Brazil. Although this treatment arm is related to
the previous ones, it addresses not the efficiency of the institutions,
but their legitimacy \citep{nivette2016institutional}. We believe that
the demand for harsher laws is on the rise in Brazil, and this may be a
major reason why citizens justify their support for mob violence. In
fact, Brazilians are often vocal about their preference for repressive
legal punishment. In a recent article in \emph{The Wall Street Journal},
a bar owner justified the lynching of the local thug who killed his son
by saying that ``\emph{even if he had been put behind bars for 100 years
it wouldn't have been enough to pay for all his crimes}''
\citep{pearson2018latam}. We hypothesise that many Brazilians also share
this view.

The experiment consists of three treatment conditions and one control
group. Respondents will read an excerpt of a news article describing a
real lynching case. We have slightly edited the original text so that
respondents have no prior knowledge of the crime.\footnote{The original
  article is available at the following address:
  \url{https://jr.jor.br/2020/05/01/homem-e-linchado-na-vila-progresso}.
  Access: August 2020.} The vignette for the control group includes no
information about the reasons behind the lynching. We will ask
respondents to show their level of support for mob violence using a
0-100 slider, where 0 means no support and 100 means full support.
Respondents in each of the three treatment arms will read the same
piece, but with one additional sentence explaining the motivations
behind the lynching. The vignettes are as follows:

\begin{itemize}
\tightlist
\item
  \emph{Control group}: A man was lynched last Friday in Jundiaí, São
  Paulo. According to the neighbours, he tried to break into a house but
  was immobilised and beaten by members of the community.\footnote{In
    Portuguese: Um homem foi linchado na última sexta-feira em Jundiaí,
    São Paulo. De acordo com vizinhos, ele tentou invadir uma residência
    mas foi imobilizado e agredido por membros da comunidade.}
\end{itemize}

\begin{itemize}
\tightlist
\item
  \emph{Treatment 01 - Police ineffectiveness}: A man was lynched last
  Friday in Jundiaí, São Paulo. According to the neighbours, he tried to
  break into a house but was immobilised and beaten by members of the
  community. \textbf{One of the residents who took part in the lynching
  said they had beaten the suspect because ``the police never patrols
  the area''}.\footnote{In Portuguese: Um homem foi linchado na última
    sexta-feira em Jundiaí, São Paulo. De acordo com vizinhos, ele
    tentou invadir uma residência mas foi imobilizado e agredido por
    membros da comunidade. \textbf{Um dos moradores envolvidos no
    linchamento disse que eles agrediram o suspeito porque ``a polícia
    nunca patrulha o local''.}}
\end{itemize}

\begin{itemize}
\tightlist
\item
  \emph{Treatment 02 - Criminal justice ineffectiveness}: A man was
  lynched last Friday in Jundiaí, São Paulo. According to the
  neighbours, he tried to break into a house but was immobilised and
  beaten by members of the community. \textbf{One of the residents who
  took part in the lynching said they had beaten the suspect because
  ``the judicial system is too slow and the perpetrator is on the street
  until the case is heard''}.\footnote{In Portuguese: Um homem foi
    linchado na última sexta-feira em Jundiaí, São Paulo. De acordo com
    vizinhos, ele tentou invadir uma residência mas foi imobilizado e
    agredido por membros da comunidade. \textbf{Um dos moradores
    envolvidos no linchamento disse que eles agrediram o suspeito porque
    ``a justiça é muito lenta e os criminosos ficam soltos até o
    julgamento''}.}
\end{itemize}

\begin{itemize}
\tightlist
\item
  \emph{Treatment 03 - Demand for harsher legal punishment}: A man was
  lynched last Friday in Jundiaí, São Paulo. According to the
  neighbours, he tried to break into a house but was immobilised and
  beaten by members of the community. \textbf{One of the residents who
  took part in the lynching said they had beaten the suspect because
  ``the judicial punishment is not harsh enough''}.\footnote{In
    Portuguese: Um homem foi linchado na última sexta-feira em Jundiaí,
    São Paulo. De acordo com vizinhos, ele tentou invadir uma residência
    mas foi imobilizado e agredido por membros da comunidade. \textbf{Um
    dos moradores envolvidos no linchamento disse que eles agrediram o
    suspeito porque ``a punição da justiça não é dura o suficiente''}.}
\end{itemize}

@@

\begin{itemize}
\tightlist
\item
  You will be shown a news article. Please read it carefully. After you
  read the article, we will ask you one question about it.\footnote{In
    Portuguese: Uma notícia será apresentada para você. Por favor, leia
    a notícia com atenção. Após você ler o artigo, faremos uma pergunta
    sobre ele.}
\end{itemize}

@@

\begin{itemize}
\tightlist
\item
  Would you think that the lynching was justified? Please use the slider
  below to indicate your opinion. For disagreement, use 0 to 49; for
  agreement, use 51 to 100. Please use 50 if you neither agree nor
  disagree.\footnote{In Portuguese: Você acha que o linchamento foi
    correto? Por favor, use a barra abaixo para indicar sua opinião.
    Para discordar, use de 0 a 49; para concordar, use de 51 a 100. Por
    favor, use 50 para não concordar nem discordar.}
\end{itemize}

\hypertarget{sample-1}{%
\subsection{Sample}\label{sample-1}}

@@

\hypertarget{estimation-1}{%
\subsection{Estimation}\label{estimation-1}}

@@

% More bibliography
\bibliography{references.bib}
